---
title: "Case Study 1 by Data Junkies"
author: "Team Data Junkies - Chris Roche, Puri Rudick, and Ryan Kinney"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---
The study of beers and breweries in the United States was conducted using over 2000 beers from more than 500 breweries.

Some of the focuses of this exploratory data analysis were placed on International Bitterness Units (IBU), Alcohol by volume (ABV), and regional differences with respect to states.

This analysis is useful for Budweiser future business decision.
```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(stringr)
library(usmap)
library(ggnewscale)
library(caret)
library(class)
library(Rmisc)
library(gridExtra)
library(ggpubr)
```

### Read in the data from csv files
```{r}
beerDataOrig = read.csv("Beers.csv",header = TRUE)
breweryDataOrig = read.csv("Breweries.csv",header = TRUE)
```

### 1. How many breweries are present in each state?

The number of breweries in each state as shown below.
The darker color represents the higher number of breweries with number labeled.

The plot shows that Colorado has the most breweries in the country, which is 47 breweries.
```{r}
breweryDataOrig = as.data.frame(breweryDataOrig)

Breweries.in.each.state = breweryDataOrig %>% group_by(State) %>% tally()
names(Breweries.in.each.state)[1] <- "state"
Breweries.in.each.state$state <-
  str_replace_all(string=Breweries.in.each.state$state, 
                  pattern=" ", repl="")

centroid_labels <- 
  utils::read.csv(system.file("extdata",
                              paste0("us_", "states", "_centroids.csv"), 
                              package = "usmap"), stringsAsFactors = FALSE)

labels <- 
  merge(x = centroid_labels, 
        y = Breweries.in.each.state, 
        by.x = "abbr", 
        by.y = "state", 
        all.x=TRUE)

names(labels)[6] <- "breweries_count"

plot_usmap(data = Breweries.in.each.state, values = "n", labels=FALSE) + 
  scale_fill_continuous(low = "white", 
                        high = "orange", 
                        name = "# of Breweries", 
                        label = scales::comma) + 
  theme(legend.position = "right") + 
  theme(panel.background = element_rect(colour = "black")) + 
  labs(title = "Number of Breweries by State") + 
  geom_text(data = labels, 
            ggplot2::aes(x = x, y = y, 
            label = scales::number(breweries_count,accuracy = 1)),
            color = "black")
```

Quick Facts:

- Colorado has the most breweries with 47
- West Virginia, North Dakota, and South Dakota each have one brewery
- Missouri (the home of Anheuser-Busch & Budweiser) has 9 breweries (Ranks 20th)


### 2. Merge beer data with the breweries data.  Print the first 6 observations and the last six observations to check the merged file.

Beer and breweries data are merged using Brewery_ID as an index.

Below are first and last six observations on the merged beer dataset.
```{r}
mergedBeerData <- merge(x = beerDataOrig, y = breweryDataOrig,
                  by.x = "Brewery_id", by.y = "Brew_ID", all = TRUE)

names(mergedBeerData)[names(mergedBeerData)=="Name.x"] <- "Beer_name"
names(mergedBeerData)[names(mergedBeerData)=="Name.y"] <- "Brewery_name"

head(mergedBeerData)
tail(mergedBeerData)
```

### 3. Address the missing values in each column.

There are missing/empty values in three columns:

- ABV: 62 values missing
- IBU: 1,005 values missing, and
- Sty: 5 values missing

Totaling in 1,007 rows with missing value.
```{r}
mergedBeerData$Style <- 
  lapply(mergedBeerData$Style,
         function(x) if(identical(x,"")) NA else x)


sapply(mergedBeerData, function(x) sum(is.na(x)))

row.has.na <- apply(mergedBeerData, 1, function(x){any(is.na(x))})
sum(row.has.na)
```

After filtering the missing/empty values out of the merged beer dataset, there are 1,405 rows of data in total that we work with.
```{r}
# Filtered Merged Data
mergedBeerData.filtered <- mergedBeerData[!row.has.na,]
nrow(mergedBeerData.filtered)

mergedBeerData.ABV_filtered = mergedBeerData %>% filter(!is.na(ABV))
nrow(mergedBeerData.ABV_filtered)

mergedBeerData.IBU_filtered = mergedBeerData %>% filter(!is.na(IBU))
nrow(mergedBeerData.IBU_filtered)
```

### 4.Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

The median of Alcohol By Volume (ABV) and International Bitterness Units (IBU) for each state as shown below.

From the bar charts, Maine has the highest number for both ABV (0.067 or 6.7%) and IBU (61 unit).
```{r message = FALSE, warning = FALSE}
# All rows NA filtered data
medianData <- aggregate(mergedBeerData.filtered[, 4:5],
                        list(mergedBeerData.filtered$State), median)
names(medianData)[names(medianData)=="Group.1"] <- "State"

# All ABV NA rows filtered
median.ABV <- aggregate(mergedBeerData.ABV_filtered[, 4],
                        list(mergedBeerData.ABV_filtered$State), median)
names(median.ABV)[names(median.ABV)=="Group.1"] <- "State"
names(median.ABV)[names(median.ABV)=="x"] <- "ABV"

# All IBU NA rows filtered
median.IBU <- aggregate(mergedBeerData.IBU_filtered[, 5],
                        list(mergedBeerData.IBU_filtered$State), median)
names(median.IBU)[names(median.IBU)=="Group.1"] <- "State"
names(median.IBU)[names(median.IBU)=="x"] <- "IBU"

# Merged data for ABV NA rows filtered and IBU NA rows filtered
median.ABV.IBU_filtered <- merge(x = median.ABV, y = median.IBU,
                                 by = "State", all = TRUE)
median.ABV.IBU_filtered$IBU[is.na(median.ABV.IBU_filtered$IBU)] <- 0

g.mid <- ggplot(median.ABV.IBU_filtered,aes(x=1, y = reorder(State,ABV))) +
  geom_text(aes(label=State)) +
  geom_segment(aes(x=0.94,xend=0.96,yend=State)) +
  geom_segment(aes(x=1.04,xend=1.062,yend=State)) +
  ggtitle("") +
  ylab(NULL) +
  scale_x_continuous(expand=c(0,0),limits=c(0.94,1.062)) +
  theme(axis.title=element_blank(),
        panel.grid=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.background=element_blank(),
        axis.text.x=element_text(color=NA),
        axis.ticks.x=element_line(color=NA),
        plot.margin = unit(c(1,-1,1,-1), "mm"))

g1 <- ggplot(data = median.ABV.IBU_filtered, aes(x = reorder(State,ABV), y = ABV, label = ABV)) +
  geom_bar(stat = "identity", fill = "sandybrown") + 
  ggtitle("Alcohol By Volume (ABV)") +
  geom_text(aes(y = ABV+0.004), 
            size = 3,
            vjust = .5) +
  theme(panel.background = element_blank(),
        axis.title.x = element_blank(), 
        axis.text.x=element_text(color=NA),
        axis.ticks.x=element_line(color=NA),
        axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        legend.position = "none",
        plot.title = element_text(hjust = 1),
        plot.margin = unit(c(1,0,1,0), "mm")) +
  scale_y_reverse() +
  coord_flip(ylim = c(0.075,0.03))

g2 <- ggplot(data = median.ABV.IBU_filtered, aes(x = reorder(State,ABV), y = IBU, fill = IBU, label=ifelse(IBU < 1, "NA", IBU))) +
  xlab(NULL) +
  geom_bar(stat = "identity") +
  ggtitle("International Bitterness Units (IBU)") +
  geom_text(aes(y = IBU+4),
            size = 3,
            vjust = .5) +
  scale_fill_gradient2(low = "grey25", 
                       high = "black",
                       name = "IBU") +
  theme(panel.background = element_blank(),
        axis.title.x = element_blank(), 
        axis.text.x=element_text(color=NA),
        axis.ticks.x=element_line(color=NA),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(),
        legend.position = "none",
        plot.margin = unit(c(1,5,1,0), "mm")) +
  coord_flip(ylim = c(0,65))

gg1 <- ggplot_gtable(ggplot_build(g1))
gg2 <- ggplot_gtable(ggplot_build(g2))
gg.mid <- ggplot_gtable(ggplot_build(g.mid))

grid.arrange(gg1,gg.mid,gg2,ncol=3,widths=c(4/9,1/9,4/9))
```

Combined the two bar charts together with radial stacked bar chart.
```{r}
median.ABV.IBU_filtered %>%
  ggplot(aes(x = reorder(State,ABV))) +
  geom_col(aes(y = ABV*1000, fill = ABV*100)) +
  scale_fill_gradient2(low = "white", 
                       high = "orange",
                       name = "Median ABV") +
  geom_text(aes(y = ABV*1000-8, label = round(ABV*100, 3)), color = "chocolate4", size = 3) +
  new_scale_fill() +
  geom_col(aes(y = -IBU, fill = IBU)) +
  scale_fill_gradient2(low = "grey25", 
                       high = "black",
                       name = "Median IBU") +
  geom_text(aes(y = -6, label = round(IBU, 0)), color = "white", size = 3) +
  geom_text(aes(y = ABV*1000+10, label = State)) + 
  scale_y_continuous(limits = c(-70, 80)) +
  coord_polar() +
  theme_void() +
  labs(title = "Median of Alcohol By Volume (ABV) and\nInternational Bitterness Units (IBU) for Each State",
       subtitle = "   * The bar chart ordered clockwise by median ABV value\n   ** The darker color represents higher value of median ABV and IBU")
```

### 5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?

On average:
- Colorado and Kentucky has the most alcoholic (6.25%) 
- Maine has the most bitter beer (61 IBU)
```{r}
# arrange by ABV in descending order, print state for first observation
# arrange by ABV in descending order, print state for first observation
median.ABV.IBU_filtered$ABV.rank <- dense_rank(-median.ABV.IBU_filtered$ABV)
with(median.ABV.IBU_filtered,
     median.ABV.IBU_filtered$State[median.ABV.IBU_filtered$ABV.rank == 1])

median.ABV.IBU_filtered$IBU.rank <- dense_rank(-median.ABV.IBU_filtered$IBU)
with(median.ABV.IBU_filtered,
     median.ABV.IBU_filtered$State[median.ABV.IBU_filtered$IBU.rank == 1])
```

By individual beers:
- Colorado has the highest ABV beer at 12.8%, Lee Hill Series Vol. 5
by Upslope Brewing Company
- Oregon has the most bitter beer at an IBU of 138, Bitter B**** Imperial IPA 
by Astoria Brewing Company
```{r}
# arrange by ABV in descending order, print state for first observation
mergedBeerData$ABV.rank <- dense_rank(-mergedBeerData$ABV)
filter(mergedBeerData,ABV.rank == 1)

mergedBeerData$IBU.rank <- dense_rank(-mergedBeerData$IBU)
filter(mergedBeerData,IBU.rank == 1)
```

### 6. Comment on the summary statistics and distribution of the ABV variable.

- The range of ABV’s is from a minimum of 0.1% to a maximum of 12.8%
- The average ABV is 6% with a standard deviation of 1.4%
```{r}
mergedBeerData.ABV_filtered %>%
    summarize(mean=mean(ABV), 
              sd=sd(ABV), 
              median=median(ABV), 
              range=max(ABV)-min(ABV), 
              min(ABV), 
              max(ABV), 
              IQR=IQR(ABV))

mergedBeerData.ABV_filtered %>% 
  ggplot(aes(x=ABV)) + 
  geom_histogram(binwidth = 0.004, color = "black", fill = "darkorange2") +
  labs(title = "Alcohol by Volume (ABV)")
```

- The range of IBU’s is from a minimum of 4 to a maximum of 138 unit
- The average IBU is 43 unit with a standard deviation of 26
```{r}
mergedBeerData.IBU_filtered %>%
    summarize(mean=mean(IBU), 
              sd=sd(IBU), 
              median=median(IBU), 
              range=max(IBU)-min(IBU), 
              min(IBU), 
              max(IBU), 
              IQR=IQR(IBU))

mergedBeerData.IBU_filtered %>% 
  ggplot(aes(x=IBU)) + 
  geom_histogram(binwidth = 5, color = "black", fill = "darkorange2") +
  labs(title = "International Bitterness Units (IBU)")

mergedBeerData.filtered %>%
    summarize(mean=mean(ABV), 
              sd=sd(ABV), 
              median=median(ABV), 
              range=max(ABV)-min(ABV), 
              min(ABV), 
              max(ABV), 
              IQR=IQR(ABV))
```

### 7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.

From the scatter plot with a regression line, there is a positive correlation between a beer’s ABV and its IBU.  There is strong evidence at the alpha = .05 level of significance to suggest that there is a linearly correlated relationship between a beer’s ABV and its IBU (p-value <.0001), with Pearson's Correlation Coefficient = 0.67 and r-square = 0.45.


```{r message = FALSE, warning = FALSE}
mergedBeerData.filtered %>%
  ggplot(aes(x = ABV,y = IBU)) + 
  geom_jitter() + 
  ggtitle("ABV vs. IBU") + 
  geom_smooth(method="lm", se=F) + 
  stat_cor(label.y = 150, 
           aes(label = paste(..rr.label.., 
                             ..p.label.., 
                             sep = "~`,`~"))) +
  stat_regline_equation(label.y = 145) +
  theme_bw()

cor.test(mergedBeerData.filtered$ABV,mergedBeerData.filtered$IBU)
```

### 8. Use KNN classification to investigate the relationship between IBU and ABV in Ales and IPAs. 

Key results of the Confusion Matrix:

- Accuracy : 0.8232
- 95% CI : (0.7597, 0.8758)
- Sensitivity : 0.8558
- Specificity : 0.7792 

```{r message = FALSE, warning = FALSE}
mergedBeerData.filtered$AleStyle <- 
  ifelse(grepl("IPA", mergedBeerData.filtered$Style, fixed = TRUE),"IPA",
         ifelse(grepl("Ale", mergedBeerData.filtered$Style, fixed = TRUE),"Ale","Other"))

mergedBeerData.filtered$AleByName <- 
  ifelse(grepl("India Pale Ale", mergedBeerData.filtered$Beer_name, fixed = TRUE),"IPA",
         ifelse(grepl("IPA", mergedBeerData.filtered$Beer_name, fixed = TRUE),"IPA",
                ifelse(grepl("Ale", mergedBeerData.filtered$Beer_name, fixed = TRUE),"Ale","Other")))

aleBeerFiltered <- mergedBeerData.filtered %>% filter(AleByName == "IPA" | AleByName == "Ale")

set.seed(4)
splitPerc = .7

trainIndices = sample(1:dim(aleBeerFiltered)[1], round(splitPerc * dim(aleBeerFiltered)[1]))
beerTrain = aleBeerFiltered[trainIndices,]
beerTest = aleBeerFiltered[-trainIndices,]

# run knn
classifications = knn(beerTrain[,4:5], beerTest[,4:5], beerTrain$AleByName, k = 5, prob = TRUE)
summary(classifications)

# Confusion Matrix
table(beerTest$AleByName, classifications)
CM = confusionMatrix(table(beerTest$AleByName, classifications))
CM

```

### 9. More useful information: Expanding KNN to detect non-Ale or IPA (other) beers. 
```{r}
set.seed(4)
splitPerc = .7

trainIndices.allType = sample(1:dim(mergedBeerData.filtered)[1], round(splitPerc * dim(mergedBeerData.filtered)[1]))
beerTrain.allType = mergedBeerData.filtered[trainIndices.allType,]
beerTest.allType = mergedBeerData.filtered[-trainIndices.allType,]

# run knn
classifications = knn(beerTrain.allType[,4:5], beerTest.allType[,4:5], beerTrain.allType$AleByName, k = 5, prob = TRUE)
summary(classifications)

# Confusion Matrix
table(beerTest.allType$AleByName, classifications)
CM.allType = confusionMatrix(table(beerTest.allType$AleByName, classifications))
CM.allType
```


### 9. More useful information: Median IBU by State
```{r}
median.ABV.IBU_filtered = as.data.frame(median.ABV.IBU_filtered)

names(median.ABV.IBU_filtered)[1] <- "state"
median.ABV.IBU_filtered$state <-
  str_replace_all(string=median.ABV.IBU_filtered$state, 
                  pattern=" ", repl="")

centroid_labels <- 
  utils::read.csv(system.file("extdata",
                              paste0("us_", "states", "_centroids.csv"), 
                              package = "usmap"), stringsAsFactors = FALSE)

labels <- 
  merge(x = centroid_labels, 
        y = median.ABV.IBU_filtered, 
        by.x = "abbr", 
        by.y = "state", 
        all.x=TRUE)

plot_usmap(data = median.ABV.IBU_filtered, values = "IBU", labels=FALSE) + 
  scale_fill_continuous(low = "white", 
                        high = "darkgreen", 
                        name = "Median IBU", 
                        label = scales::comma) + 
  theme(legend.position = "right") + 
  theme(panel.background = element_rect(colour = "black")) + 
  labs(title = "Median IBU by State")

```

Observation: There is a popular style of beer called a "West Coast IPA", but the West Coast does not have more bitter beer than other regions on average.



### 9. More useful information: Median ABV by State
```{r}
plot_usmap(data = median.ABV.IBU_filtered, values = "ABV", labels=FALSE) + 
  scale_fill_continuous(low = "white", 
                        high = "darkblue", 
                        name = "Median ABV", 
                        label = scales::comma) + 
  theme(legend.position = "right") + 
  theme(panel.background = element_rect(colour = "black")) + 
  labs(title = "Median ABV by State")
```

Observation: Utah and Arkansas had laws until 2020 limiting the alcohol content of beer.



### 9. More useful information: Beer Size (Ounces) by State

There are 9 states that the biggest beer container size is only 12 ounces
- Alaska
- Alabama
- Arkansas
- DC
- Georgia
- North Dakota
- South Dakota
- Tennessee
- West Virginia

It might be a good opportunity for Budweiser to launch a bigger size product in these states.
```{r}
max.size <- aggregate(mergedBeerData[, 7],
                      list(mergedBeerData$State), max)
names(max.size)[names(max.size)=="Group.1"] <- "state"
names(max.size)[names(max.size)=="x"] <- "max.Ounces"

max.size$max.bin = cut(max.size$max.Ounces, 
                       breaks = c(0,11,12,16,32), 
                       labels = c("< 12 Ounces","12 Ounces","16 Ounce", ">16 Ounce"))

filter(max.size,max.bin == "12 Ounces")

max.size$state <-
  str_replace_all(string=max.size$state, 
                  pattern=" ", repl="")

centroid_labels <- 
  utils::read.csv(system.file("extdata",
                              paste0("us_", "states", "_centroids.csv"), 
                              package = "usmap"), stringsAsFactors = FALSE)

labels <- 
  merge(x = centroid_labels, 
        y = max.size, 
        by.x = "abbr", 
        by.y = "state", 
        all.x=TRUE)

plot_usmap(data = max.size, values = "max.bin", labels=TRUE) + 
  scale_fill_manual(values = c("white", "skyblue1", "skyblue4"), 
                    name = "Max Container Size") +
  theme(legend.position = "right") + 
  theme(panel.background = element_rect(colour = "black"),
        legend.text = element_text(size = 10)) + 
  labs(title = "Maximum Beer Container Size by State")
```
